---
title: "PA1_template"
author: "Ross Voorhees"
date: "Saturday, December 19, 2015"
output: html_document
keep_md: yes
---



```{r, echo=FALSE, message=F, warning=F}
library(plyr)
library(dplyr)
library(sqldf)
library(lubridate)
library(ggplot2)
library(scales)
library(timeDate)
library(memisc, warn.conflicts = FALSE, quietly=TRUE)
```

###Loading and preprocessing the data


######Creates Directory if doesn't Exist

```{r}
if(!file.exists("./Activities")) {dir.create("./Activities")}
```

######Sets Vector to Wearables

```{r}
dir <- "wearables"
```
######Downloads DataSet and Unzips

```{r}
fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileurl,destfile="Activities\\ActivitiesData.zip")
unzip("Activities\\ActivitiesData.zip", list = FALSE,exdir="Activities", overwrite = TRUE)
```




######Read Files into vector

```{r}
Activities <- read.csv(file="activity.csv",header=TRUE, sep=,)
```


######Remove Missing Values & Zero Values

```{r}
Activities <- na.omit(Activities)
```

######Converts date to Date format and merges into vector

```{r}
Activities$date <- as.Date(Activities$date, format = "%Y-%m-%d")
```

###Calculate the total number of steps taken per day



```{r message=F, warning=F}
summarise(group_by(Activities),TotalSteps=sum(steps, na.rm=TRUE))
```



###Make a histogram of the total number of steps taken each day

```{r}
astep <- summarise(group_by(Activities, date),TotalSteps=sum(steps, na.rm=TRUE))
ggplot(astep, aes(TotalSteps)) + geom_histogram(fill="Blue")+ggtitle("Steps Taken Per Day")+ylab("Number of Days")
```       
    
###Calculate and report the mean and median of the total number of steps taken per day

```{r}
summarise(astep,mean=mean(TotalSteps), median=median(TotalSteps) )
```


###What is the average daily activity pattern?

```{r}
istep <- summarise(group_by(Activities,interval), avgsteps=mean(steps))
plot(istep$interval, istep$avgsteps, type="l", xlab="5 minute Interval", ylab="Avg Steps")
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r message=F, warning=F}
istepm <- sqldf("select interval, avgsteps Steps from istep order by avgsteps desc")
head(istepm, 1)
```

###Imputing missing values

```{r}
Activities2 <- read.csv(file="activity.csv",header=TRUE, sep=,)
```



###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r,}
sum(is.na(Activities2$steps))
```

    Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated.

    This finds the mean and median for each interval.

```{r}
m <- summarise(group_by(Activities, interval),mean=mean(steps), median=median(steps) )
```

    Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
f <- sqldf("Select a.date, a.interval, CASE WHEN a.steps is null THEN m.median ELSE a.steps END steps
          from Activities2 a Left Outer Join m on a.interval = m.interval ")
```

###Make a histogram of the total number of steps taken each day

```{r }
bstep <- summarise(group_by(f, date),TotalSteps=sum(steps, na.rm=TRUE))

ggplot(bstep, aes(TotalSteps)) + geom_histogram(fill="Blue")+ggtitle("Steps Taken Per Day")+ylab("Number of Days")
```

    Mean and Median After NA Replacement

```{r}
summarise(bstep,mean=mean(TotalSteps), median=median(TotalSteps) )
```

    Mean and Median Before NA Replacement
    
    Yes the mean decreased, but the median increased
```{r}
summarise(astep,mean=mean(TotalSteps), median=median(TotalSteps) )
```
    The mean decreased, but the median increased

###Are there differences in activity patterns between weekdays and weekends?

    Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or                                 weekend day.

```{r}
f$date <- as.Date(f$date, format = "%Y-%m-%d")
f$weekday <- weekdays(f$date)
f$weekendflag <- isWeekend(f$date)
f$weekendflag[f$weekendflag == "TRUE"] <- "Weekend"
f$weekendflag[f$weekendflag == "FALSE"] <- "Weekday"
```

    Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and 
    the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
w <- summarise(group_by(f,interval,weekendflag), NumSteps=mean(steps))

ggplot(w, aes(interval,NumSteps)) + geom_line(fill="Blue")+ylab("Number of Steps")+facet_grid(weekendflag ~ .)
```



















